---
title: "Graphing with R"
author: "Yirui-Irena-Xu"
output: 
  html_document:
    highlight: zenburn
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE,  echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
library(infer)
```

# Excess rentals in TfL bike sharing

We can get the latest TfL data on how many bikes were hired every single day by running the following

```{r, get_tfl_data, cache=TRUE}
url <- "https://data.london.gov.uk/download/number-bicycle-hires/ac29363e-e0cb-47cc-a97a-e216d900a6b0/tfl-daily-cycle-hires.xlsx"
```

```{r, clean_the_date}
# Use read_excel to read it as dataframe
bike0 <- read_excel(bike.temp,
                   sheet = "Data",
                   range = cell_cols("A:B"))

# change dates to get year, month, and week
bike <- bike0 %>% 
  clean_names() %>% 
  rename (bikes_hired = number_of_bicycle_hires) %>% 
  mutate (year = year(day),
          month = lubridate::month(day, label = TRUE),
          week = isoweek(day))
```



We can visualize how actual rentals varied from expectations based on the data. 

```{r bike_graph1, fig.width=10}
bike_graph1 <- bike %>% 
  filter(year>=2015) %>% 
  group_by(month) %>% 
  mutate(expected_rentals=median(bikes_hired)) %>% 
  ungroup %>% 
  group_by(month, year) %>% 
  summarise(expected_rentals = median(expected_rentals), 
            actual_rentals = median(bikes_hired)) %>% 
  mutate(excess_rentals = actual_rentals - expected_rentals)
  
ggplot(bike_graph1, 
       aes(x=month, group=1))+
  geom_ribbon(aes(ymin = ifelse(actual_rentals < expected_rentals, 
                                actual_rentals, expected_rentals), 
                  ymax = expected_rentals),
              fill= "green", 
              alpha=0.1)+
  geom_ribbon(aes(ymin=expected_rentals, 
                  ymax=ifelse(actual_rentals > expected_rentals, 
                              actual_rentals, expected_rentals)),
              fill="red", 
              alpha=0.1)+
    geom_line(aes(y=expected_rentals), 
              color= "blue", 
              size=0.5)+
    geom_line(aes(y=actual_rentals))+
    facet_wrap(~year)+
  theme_bw()
  
```

Besides, we can looks at percentage changes from the expected level of weekly rentals. The two grey shaded rectangles correspond to the second (weeks 14-26) and fourth (weeks 40-52) quarters.

```{r bike_graph2, fig.width=10}
bike_graph2 <- bike %>% 
  filter(year>=2015) %>% 
  group_by(week) %>% 
  mutate(weekly_average = median(bikes_hired)) %>% 
  ungroup %>% 
  group_by(week, year) %>% 
  summarise(weekly_average = mean(weekly_average), 
            actual_bikes_hired = median(bikes_hired)) %>% 
  mutate(percentage_change = actual_bikes_hired / weekly_average - 1)

 
ggplot(bike_graph2, 
       aes(x=week, group=1))+
  geom_rect(xmin=13,xmax=26, 
            ymin=-0.6, ymax=0.6, 
            colour="grey", 
            alpha=0.003)+
  geom_rect(xmin=39,xmax=52, 
            ymin=-0.6,ymax=0.6, 
            colour="grey", 
            alpha=0.003)+
  geom_ribbon(aes(ymin=0,
                  ymax=ifelse(percentage_change>0,
                              percentage_change ,0)), 
              fill="green" ,
              alpha=0.15)+
  geom_ribbon(aes(ymin=ifelse(percentage_change<0, 
                              percentage_change,0), 
                  ymax=0), 
              fill="red", 
              alpha=0.15)+
  geom_line(aes(y=percentage_change))+
  geom_rug(side="week", 
           aes(color=ifelse(percentage_change<0,
                            "red", "green")))+
  guides(color=FALSE)+
  scale_x_continuous(breaks=c(13,26,39,52))+
  scale_y_continuous(labels=scales::percent)+
  facet_wrap(~year)+
  theme_bw()

```